---
title: "Cosmos Birth Vaccines"
date: "`r format(Sys.Date(), '%Y-%m-%d')`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

library(dplyr)
library(ggplot2)
library(tigris)
library(sf)
library(patchwork)
library(vroom)
library(ggrepel)

options(tigris_use_cache = TRUE)
if (!dir.exists("maps")) dir.create("maps")
```

Comparing two Epic Cosmos population bases:

- **All patients** — `standard/data.csv.gz`: all patients with birthing parent
  information, no additional filter.
- **CPT birth cohort** — `standard/data_cpt_birth.csv.gz`: subset filtered to a billed
  procedure indicating birth (CPT 99460-99465, 99468, 99477).



```{r load-data}
all_patients <- vroom::vroom("standard/data.csv.gz", show_col_types = FALSE) %>%
  filter(geography != "00") %>%
  mutate(date = as.Date(time, "%m-%d-%Y"))
cpt_birth <- vroom::vroom("standard/data_cpt_birth.csv.gz", show_col_types = FALSE) %>%
  filter(geography != "00") %>%
  mutate(date = as.Date(time, "%m-%d-%Y"))

# Snapshot at the latest year both exports cover
common_date <- max(intersect(all_patients$date, cpt_birth$date)) %>%
  as.Date(origin = "1970-01-01")
time_label <- format(common_date, "%Y")

all_snap <- all_patients %>% filter(date == common_date)
cpt_snap <- cpt_birth     %>% filter(date == common_date)

states_sf <- tigris::states(cb = TRUE, year = 2020) %>%
  filter(!STATEFP %in% c("60", "66", "69", "72", "78")) %>% # drop territories
  select(geography = STATEFP, state_name = NAME, geometry) %>%
  tigris::shift_geometry() # inset AK/HI below the continental US

map_all <- states_sf %>% left_join(all_snap, by = "geography")
map_cpt <- states_sf %>% left_join(cpt_snap, by = "geography")

pct_measures <- list(
  list(col = "epic_pct_vitamin_k", label = "Vitamin K", slug = "vitamin_k"),
  list(col = "epic_pct_rsv",       label = "RSV immunization", slug = "rsv"),
  list(col = "epic_pct_hepb",      label = "HepB vaccination", slug = "hepb")
)

diff_data <- all_snap %>%
  select(geography, all_of(vapply(pct_measures, `[[`, character(1), "col"))) %>%
  rename_with(~ paste0(., "_all"), -geography) %>%
  full_join(
    cpt_snap %>%
      select(geography, all_of(vapply(pct_measures, `[[`, character(1), "col")), epic_n_patients) %>%
      rename_with(~ paste0(., "_cpt"), -geography),
    by = "geography"
  ) %>%
  left_join(
    all_snap %>% select(geography, epic_n_patients_all = epic_n_patients),
    by = "geography"
  ) %>%
  left_join(states_sf %>% st_drop_geometry() %>% select(geography, state_name), by = "geography") %>%
  mutate(state_abb = state.abb[match(state_name, state.name)])

for (m in pct_measures) {
  diff_data[[paste0("diff_", m$slug)]] <-
    diff_data[[paste0(m$col, "_cpt")]] - diff_data[[paste0(m$col, "_all")]]
}
diff_data$ratio_n <- diff_data$epic_n_patients_cpt / diff_data$epic_n_patients_all

map_diff <- states_sf %>% left_join(diff_data, by = "geography")
```

```{r helpers}
state_panel <- function(map_df, fill_var, title, legend_name = "", diverging = FALSE) {
  if (diverging) {
    lim   <- max(abs(map_df[[fill_var]]), na.rm = TRUE)
    scale <- scale_fill_distiller(
      name = legend_name, palette = "RdBu", direction = 1,
      limits = c(-lim, lim), na.value = "grey85"
    )
  } else {
    scale <- scale_fill_viridis_c(
      name = legend_name, option = "plasma", na.value = "grey85",
      breaks = scales::breaks_pretty(n = 4)
    )
  }

  ggplot(map_df) +
    geom_sf(aes(fill = .data[[fill_var]]), color = "white", linewidth = 0.1) +
    scale +
    labs(title = title) +
    theme_void(base_size = 10) +
    theme(
      plot.title       = element_text(face = "bold", size = 11, hjust = 0.5),
      legend.position  = "bottom",
      legend.key.width = unit(1.2, "cm"),
      legend.title     = element_text(size = 8),
      legend.text      = element_text(size = 7)
    )
}

# Builds and saves the triptych (All | CPT birth | Difference) for one measure,
# returning the combined patchwork so the calling chunk can print it inline.
render_triptych <- function(m) {
  p1 <- state_panel(map_all,  m$col, "All patients", "%")
  p2 <- state_panel(map_cpt,  m$col, "CPT birth cohort", "%")
  p3 <- state_panel(map_diff, paste0("diff_", m$slug), "Difference (CPT birth − All)", "pp", diverging = TRUE)

  combined <- (p1 | p2 | p3) +
    plot_annotation(
      title    = paste0(m$label, " (%) by State — ", time_label),
      subtitle = "Epic Cosmos | All patients with birthing parent information vs. CPT birth cohort subset",
      caption  = "Grey = no data.",
      theme = theme(
        plot.title    = element_text(face = "bold", size = 14),
        plot.subtitle = element_text(color = "grey40"),
        plot.caption  = element_text(color = "grey50", size = 8)
      )
    )

  ggsave(file.path("maps", paste0("compare_", m$slug, ".png")), combined, width = 18, height = 6, dpi = 150)
  combined
}

# Builds and saves the All-vs-CPT-birth scatter plot for one measure, returning
# the ggplot so the calling chunk can print it inline.
render_scatter <- function(m) {
  x_col <- paste0(m$col, "_all")
  y_col <- paste0(m$col, "_cpt")

  scatter_data <- diff_data %>%
    filter(!is.na(.data[[x_col]]), !is.na(.data[[y_col]]))

  lim <- range(c(scatter_data[[x_col]], scatter_data[[y_col]]), na.rm = TRUE)

  cor_test <- cor.test(scatter_data[[x_col]], scatter_data[[y_col]])
  cor_label <- sprintf(
    "r = %.3f, p = %.3f", round(cor_test$estimate, 3), round(cor_test$p.value, 3)
  )

  p_scatter <- ggplot(scatter_data, aes(x = .data[[x_col]], y = .data[[y_col]])) +
    geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "grey60") +
    geom_point(color = "#1b6ca8", size = 2) +
    ggrepel::geom_text_repel(aes(label = state_abb), size = 2.8, max.overlaps = 30) +
    coord_equal(xlim = lim, ylim = lim) +
    labs(
      title    = paste0(m$label, " (%) — All vs. CPT Birth Cohort by State, ", time_label),
      subtitle = paste0(cor_label),
      x        = "All patients (%)",
      y        = "CPT birth cohort (%)"
    ) +
    theme_minimal(base_size = 11) +
    theme(plot.title = element_text(face = "bold", size = 12))

  ggsave(file.path("maps", paste0("scatter_", m$slug, ".png")), p_scatter, width = 7, height = 7, dpi = 150)
  p_scatter
}
```

## Vitamin K

```{r vitamin-k-map, fig.width=18, fig.height=6}
render_triptych(pct_measures[[1]])
```

```{r vitamin-k-scatter, fig.width=7, fig.height=7}
render_scatter(pct_measures[[1]])
```

## RSV Immunization

```{r rsv-map, fig.width=18, fig.height=6}
render_triptych(pct_measures[[2]])
```

```{r rsv-scatter, fig.width=7, fig.height=7}
render_scatter(pct_measures[[2]])
```

## HepB Vaccination

```{r hepb-map, fig.width=18, fig.height=6}
render_triptych(pct_measures[[3]])
```

```{r hepb-scatter, fig.width=7, fig.height=7}
render_scatter(pct_measures[[3]])
```

## Patient Count QC

Share of the "all patients" population that also has a CPT code indicating birth.

```{r ratio-map, fig.width=10, fig.height=8.5}
p_ratio <- ggplot(map_diff) +
  geom_sf(aes(fill = ratio_n), color = "white", linewidth = 0.1) +
  scale_fill_viridis_c(
    name = "CPT birth / All", option = "magma", na.value = "grey85",
    breaks = scales::breaks_pretty(n = 5)
  ) +
  labs(
    title    = paste0("Patient Count Ratio (CPT Birth Cohort / All) by State — ", time_label),
    subtitle = "Share of the 'all patients' population that also has a CPT code indicating birth"
  ) +
  theme_void(base_size = 11) +
  theme(
    plot.title       = element_text(face = "bold", size = 13, hjust = 0.5),
    plot.subtitle    = element_text(color = "grey40", hjust = 0.5, size = 9),
    legend.position  = "bottom",
    legend.key.width = unit(1.5, "cm")
  )

ggsave(file.path("maps", "compare_n_patients_ratio.png"), p_ratio, width = 10, height = 8.5, dpi = 150)
p_ratio
```

```{r timeseries, fig.width=9, fig.height=5}
national <- bind_rows(
  vroom::vroom("standard/data.csv.gz", show_col_types = FALSE) %>%
    filter(geography == "00") %>% mutate(source = "All patients"),
  vroom::vroom("standard/data_cpt_birth.csv.gz", show_col_types = FALSE) %>%
    filter(geography == "00") %>% mutate(source = "CPT birth cohort")
) %>%
  mutate(date = as.Date(time, "%m-%d-%Y"))

p_ts <- ggplot(national, aes(x = date, y = epic_n_patients, color = source)) +
  geom_line(linewidth = 0.9) +
  geom_point(size = 1.5) +
  scale_y_continuous(labels = scales::label_comma()) +
  scale_color_manual(values = c("All patients" = "#1b6ca8", "CPT birth cohort" = "#d1495b")) +
  labs(
    title = "National Patient Count Over Time",
    x = NULL, y = "Patients", color = NULL
  ) +
  theme_minimal(base_size = 11) +
  theme(legend.position = "bottom")

ggsave(file.path("maps", "compare_n_patients_timeseries.png"), p_ts, width = 9, height = 5, dpi = 150)
p_ts
```

