---
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE,
                       fig.width = 12, fig.height = 7, dpi = 120)

library(dplyr)
library(tidyr)
library(vroom)
library(ggplot2)
library(scales)
library(sf)
library(tigris)

options(tigris_use_cache = TRUE)

DISEASES <- c("lyme", "babesiosis", "malaria", "rmsf", "west_nile", "dengue")
DISEASE_LABELS <- c(
  lyme = "Lyme disease", babesiosis = "Babesiosis", malaria = "Malaria",
  rmsf = "RMSF", west_nile = "West Nile virus", dengue = "Dengue"
)
YEARS <- c("2023", "2024", "2025", "2026")
DROP_ST <- c("02", "15", "60", "66", "69", "72", "78")  # AK, HI, territories

dat <- vroom(
  "standard/data.csv.gz", show_col_types = FALSE,
  col_types = cols(geography = "c", time = "c")
)

for (dz in DISEASES) {
  dat[[paste0("epic_rate100k_", dz)]] <-
    dat[[paste0("epic_n_", dz)]] / dat$epic_n_patients * 1e5
}
```

# Epic Cosmos vector-borne diseases {.tabset}

## Trends and maps

```{r national-time-series}
pct_cols <- paste0("epic_pct_", DISEASES)

nat_long <- dat %>%
  filter(geography == "00") %>%
  mutate(time = as.Date(time)) %>%
  select(time, all_of(pct_cols)) %>%
  pivot_longer(-time, names_to = "measure", values_to = "value") %>%
  mutate(measure = DISEASE_LABELS[sub("^epic_pct_", "", measure)])

ggplot(nat_long, aes(time, value)) +
  geom_line(linewidth = 0.6) +
  geom_point(size = 0.8) +
  facet_wrap(~measure, scales = "free_y", ncol = 3) +
  labs(title = "National monthly percent of patients, by disease",
       x = NULL, y = "Percent of patients") +
  theme_bw(base_size = 11) +
  theme(strip.text = element_text(face = "bold"),
        axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r national-time-series-rate}
rate_cols <- paste0("epic_rate100k_", DISEASES)

nat_rate_long <- dat %>%
  filter(geography == "00") %>%
  mutate(time = as.Date(time)) %>%
  select(time, all_of(rate_cols)) %>%
  pivot_longer(-time, names_to = "measure", values_to = "value") %>%
  mutate(measure = DISEASE_LABELS[sub("^epic_rate100k_", "", measure)])

ggplot(nat_rate_long, aes(time, value)) +
  geom_line(linewidth = 0.6) +
  geom_point(size = 0.8) +
  facet_wrap(~measure, scales = "free_y", ncol = 3) +
  labs(title = "National monthly rate per 100,000 patients, by disease",
       x = NULL, y = "Cases per 100,000 patients") +
  theme_bw(base_size = 11) +
  theme(strip.text = element_text(face = "bold"),
        axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r regional-time-series}
# Five common geographic regions (National Geographic scheme; DC grouped with
# the Northeast). Regional percent = pooled counts across member states /
# pooled patients, so larger states weigh more.
REGION_STATES <- list(
  Northeast = c("CT", "DE", "DC", "ME", "MD", "MA", "NH", "NJ", "NY", "PA", "RI", "VT"),
  Southeast = c("AL", "AR", "FL", "GA", "KY", "LA", "MS", "NC", "SC", "TN", "VA", "WV"),
  Midwest   = c("IL", "IN", "IA", "KS", "MI", "MN", "MO", "NE", "ND", "OH", "SD", "WI"),
  Southwest = c("AZ", "NM", "OK", "TX"),
  West      = c("AK", "CA", "CO", "HI", "ID", "MT", "NV", "OR", "UT", "WA", "WY")
)
REGION_MAP <- stack(REGION_STATES) %>%
  transmute(state = as.character(values), region = as.character(ind))

all_fips <- vroom("../../resources/all_fips.csv.gz", show_col_types = FALSE,
                  col_types = cols(geography = "c"))

region_lookup <- all_fips %>%
  filter(nchar(geography) == 2, geography != "00") %>%
  select(geography, state) %>%
  left_join(REGION_MAP, by = "state") %>%
  filter(!is.na(region)) %>%
  select(geography, region)

n_cols_reg <- paste0("epic_n_", DISEASES)

region_long <- dat %>%
  inner_join(region_lookup, by = "geography") %>%
  mutate(time = as.Date(time)) %>%
  group_by(region, time) %>%
  summarize(across(all_of(c(n_cols_reg, "epic_n_patients")), ~ sum(.x, na.rm = TRUE)),
            .groups = "drop") %>%
  pivot_longer(all_of(n_cols_reg), names_to = "measure", values_to = "n") %>%
  mutate(value = n / epic_n_patients * 100,
         measure = DISEASE_LABELS[sub("^epic_n_", "", measure)])

ggplot(region_long, aes(time, value, color = region)) +
  geom_line(linewidth = 0.6) +
  geom_point(size = 0.8) +
  facet_wrap(~measure, scales = "free_y", ncol = 3) +
  scale_color_brewer(palette = "Dark2", name = NULL) +
  labs(title = "Regional monthly percent of patients, by disease",
       subtitle = "Five geographic regions; pooled across member states",
       x = NULL, y = "Percent of patients") +
  theme_bw(base_size = 11) +
  theme(strip.text = element_text(face = "bold"),
        axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "bottom")
```

```{r regional-time-series-log}
# Same regional monthly percent, on a log scale with one panel per disease and
# a shared y-axis so the four diseases can be compared with each other.
LOG_DISEASES <- c("lyme", "babesiosis", "malaria", "rmsf")

region_log <- region_long %>%
  filter(measure %in% DISEASE_LABELS[LOG_DISEASES]) %>%
  mutate(measure = factor(measure, levels = DISEASE_LABELS[LOG_DISEASES]))

ggplot(region_log, aes(time, value, color = region)) +
  geom_line(linewidth = 0.6) +
  geom_point(size = 0.8) +
  facet_wrap(~measure, ncol = 2) +
  scale_y_log10(labels = label_number(accuracy = 0.0001)) +
  scale_color_brewer(palette = "Dark2", name = NULL) +
  labs(title = "Regional monthly percent of patients, log scale",
       subtitle = "Lyme disease, babesiosis, malaria and RMSF; same y-axis in every panel",
       x = NULL, y = "Percent of patients (log scale)") +
  theme_bw(base_size = 11) +
  theme(strip.text = element_text(face = "bold"),
        axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "bottom")
```

```{r no-travel-comparison}
# Comparison against cosmos_vector_borne_no_travel (same six diseases, but
# restricted to patients without a travel history). Denominators differ
# between the two sessions (see that source's README), so this compares raw
# counts (epic_n_*), not percentages.
no_travel_file <- "../cosmos_vector_borne_no_travel/standard/data.csv.gz"

if (file.exists(no_travel_file)) {
  n_cols <- paste0("epic_n_", DISEASES)

  dat_all <- dat %>%
    filter(geography == "00") %>%
    mutate(time = as.Date(time), source = "All patients") %>%
    select(time, source, all_of(n_cols))

  dat_nt <- vroom(no_travel_file, show_col_types = FALSE,
                   col_types = cols(geography = "c", time = "c")) %>%
    filter(geography == "00") %>%
    mutate(time = as.Date(time), source = "No travel history") %>%
    select(time, source, all_of(n_cols))

  compare_long <- bind_rows(dat_all, dat_nt) %>%
    pivot_longer(all_of(n_cols), names_to = "measure", values_to = "value") %>%
    mutate(measure = DISEASE_LABELS[sub("^epic_n_", "", measure)])

  ggplot(compare_long, aes(time, value, color = source)) +
    geom_line(linewidth = 0.6) +
    geom_point(size = 0.8) +
    facet_wrap(~measure, scales = "free_y", ncol = 3) +
    scale_color_manual(values = c("All patients" = "#1b6ca8", "No travel history" = "#e07b39"),
                        name = NULL) +
    labs(title = "National monthly patient count, by disease — all patients vs. no travel history",
         subtitle = "Counts only (not percentages) — the two sessions use different population denominators",
         x = NULL, y = "Number of patients") +
    theme_bw(base_size = 11) +
    theme(strip.text = element_text(face = "bold"),
          axis.text.x = element_text(angle = 45, hjust = 1),
          legend.position = "bottom")
} else {
  message("cosmos_vector_borne_no_travel standard data not found; skipping comparison.")
}
```

```{r map-data}
states_sf <- tigris::states(cb = TRUE, resolution = "5m", year = 2022, progress_bar = FALSE) %>%
  filter(!STATEFP %in% DROP_ST) %>%
  transmute(geography = as.character(GEOID))

flag_cols <- paste0("epic_", DISEASES, "_suppressed_flag")

state_annual <- dat %>%
  filter(nchar(geography) == 2, geography != "00") %>%
  mutate(year = substr(time, 1, 4)) %>%
  filter(year %in% YEARS) %>%
  group_by(geography, year) %>%
  summarize(across(all_of(c(paste0("epic_pct_", DISEASES),
                             paste0("epic_rate100k_", DISEASES))),
                    ~ mean(.x, na.rm = TRUE)),
            # TRUE only when every month contributing to the annual mean was
            # suppressed for that disease - i.e. the average is entirely
            # imputed, not just partly.
            across(all_of(flag_cols), ~ mean(.x) == 1, .names = "allsupp_{.col}"),
            .groups = "drop") %>%
  mutate(year = factor(year, levels = YEARS))

draw_map <- function(col, flag_col, label, low_col = "viridis") {
  md <- states_sf %>%
    left_join(
      state_annual %>%
        select(geography, year, value = all_of(col), suppressed = all_of(flag_col)) %>%
        mutate(value = if_else(suppressed, NA_real_, value)),
      by = "geography"
    ) %>%
    st_as_sf()

  ggplot(md) +
    geom_sf(aes(fill = value), color = "white", linewidth = 0.08) +
    scale_fill_viridis_c(option = low_col, na.value = "grey70",
                         labels = label_number(accuracy = 0.01), name = NULL) +
    facet_wrap(~year, ncol = 2) +
    coord_sf(crs = 5070, datum = NA) +
    labs(title = label) +
    theme_void(base_size = 11) +
    theme(plot.title = element_text(face = "bold", hjust = 0.5, size = 13),
          strip.text = element_text(face = "bold", size = 11),
          legend.position = "right")
}
```

```{r maps, results='asis'}
for (dz in DISEASES) {
  flag_col <- paste0("allsupp_epic_", dz, "_suppressed_flag")
  print(draw_map(paste0("epic_pct_", dz), flag_col,
                 sprintf("%s — percent of patients, by year", DISEASE_LABELS[dz])))
  print(draw_map(paste0("epic_rate100k_", dz), flag_col,
                 sprintf("%s — rate per 100,000 patients, by year", DISEASE_LABELS[dz])))
}
```

## New cases

Patients with a **new** (first-time) diagnosis, from `cosmos_vector_borne_new_cases`
(SlicerDicer session 2861722). Same population base and denominator as the tab above,
but the numerator counts only first-time diagnoses, so it is a subset of the
all-diagnosis counts. Most state-month cells for the rarer diseases are suppressed
(imputed as 5), so the national and regional panels are the most reliable.

```{r nc-data}
nc_file <- "../cosmos_vector_borne_new_cases/standard/data.csv.gz"
HAVE_NC <- file.exists(nc_file)
if (!HAVE_NC) message("cosmos_vector_borne_new_cases standard data not found; skipping new-case section.")

if (HAVE_NC) {
  nc <- vroom(nc_file, show_col_types = FALSE,
              col_types = cols(geography = "c", time = "c"))
  for (dz in DISEASES) {
    nc[[paste0("epic_rate100k_new_", dz)]] <-
      nc[[paste0("epic_n_new_", dz)]] / nc$epic_n_patients * 1e5
  }
}
```

```{r nc-national, eval=HAVE_NC}

nc_nat <- nc %>%
  filter(geography == "00") %>%
  mutate(time = as.Date(time)) %>%
  select(time, all_of(paste0("epic_pct_new_", DISEASES))) %>%
  pivot_longer(-time, names_to = "measure", values_to = "value") %>%
  mutate(measure = DISEASE_LABELS[sub("^epic_pct_new_", "", measure)])

ggplot(nc_nat, aes(time, value)) +
  geom_line(linewidth = 0.6) +
  geom_point(size = 0.8) +
  facet_wrap(~measure, scales = "free_y", ncol = 3) +
  labs(title = "National monthly percent of patients with a new diagnosis, by disease",
       x = NULL, y = "Percent of patients") +
  theme_bw(base_size = 11) +
  theme(strip.text = element_text(face = "bold"),
        axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r nc-national-rate, eval=HAVE_NC}
nc_rate <- nc %>%
  filter(geography == "00") %>%
  mutate(time = as.Date(time)) %>%
  select(time, all_of(paste0("epic_rate100k_new_", DISEASES))) %>%
  pivot_longer(-time, names_to = "measure", values_to = "value") %>%
  mutate(measure = DISEASE_LABELS[sub("^epic_rate100k_new_", "", measure)])

ggplot(nc_rate, aes(time, value)) +
  geom_line(linewidth = 0.6) +
  geom_point(size = 0.8) +
  facet_wrap(~measure, scales = "free_y", ncol = 3) +
  labs(title = "National monthly rate of new diagnoses per 100,000 patients, by disease",
       x = NULL, y = "New cases per 100,000 patients") +
  theme_bw(base_size = 11) +
  theme(strip.text = element_text(face = "bold"),
        axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r nc-vs-all, eval=HAVE_NC}
# New-case counts are a subset of the all-diagnosis counts (same patients
# denominator definition), so the two series are directly comparable.
nc_n_cols <- paste0("epic_n_new_", DISEASES)
all_n_cols <- paste0("epic_n_", DISEASES)

nc_cmp <- bind_rows(
  dat %>% filter(geography == "00") %>%
    transmute(time = as.Date(time), source = "Any diagnosis",
              !!!setNames(lapply(all_n_cols, as.name), DISEASES)),
  nc %>% filter(geography == "00") %>%
    transmute(time = as.Date(time), source = "New diagnosis only",
              !!!setNames(lapply(nc_n_cols, as.name), DISEASES))
) %>%
  pivot_longer(all_of(DISEASES), names_to = "measure", values_to = "value") %>%
  mutate(measure = DISEASE_LABELS[measure])

ggplot(nc_cmp, aes(time, value, color = source)) +
  geom_line(linewidth = 0.6) +
  geom_point(size = 0.8) +
  facet_wrap(~measure, scales = "free_y", ncol = 3) +
  scale_color_manual(values = c("Any diagnosis" = "#1b6ca8",
                                 "New diagnosis only" = "#e07b39"), name = NULL) +
  labs(title = "National monthly patient count — any diagnosis vs. new diagnosis only",
       x = NULL, y = "Number of patients") +
  theme_bw(base_size = 11) +
  theme(strip.text = element_text(face = "bold"),
        axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "bottom")
```

```{r nc-regional, eval=HAVE_NC}
nc_n_reg <- paste0("epic_n_new_", DISEASES)

nc_region <- nc %>%
  inner_join(region_lookup, by = "geography") %>%
  mutate(time = as.Date(time)) %>%
  group_by(region, time) %>%
  summarize(across(all_of(c(nc_n_reg, "epic_n_patients")), ~ sum(.x, na.rm = TRUE)),
            .groups = "drop") %>%
  pivot_longer(all_of(nc_n_reg), names_to = "measure", values_to = "n") %>%
  mutate(value = n / epic_n_patients * 100,
         measure = DISEASE_LABELS[sub("^epic_n_new_", "", measure)])

ggplot(nc_region, aes(time, value, color = region)) +
  geom_line(linewidth = 0.6) +
  geom_point(size = 0.8) +
  facet_wrap(~measure, scales = "free_y", ncol = 3) +
  scale_color_brewer(palette = "Dark2", name = NULL) +
  labs(title = "Regional monthly percent of patients with a new diagnosis, by disease",
       subtitle = "Five geographic regions; pooled across member states",
       x = NULL, y = "Percent of patients") +
  theme_bw(base_size = 11) +
  theme(strip.text = element_text(face = "bold"),
        axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "bottom")
```

```{r nc-regional-log, eval=HAVE_NC}
# Same regional monthly percent on a log scale, shared y-axis (as in the
# all-diagnosis tab).
nc_region_log <- nc_region %>%
  filter(measure %in% DISEASE_LABELS[LOG_DISEASES], value > 0) %>%
  mutate(measure = factor(measure, levels = DISEASE_LABELS[LOG_DISEASES]))

ggplot(nc_region_log, aes(time, value, color = region)) +
  geom_line(linewidth = 0.6) +
  geom_point(size = 0.8) +
  facet_wrap(~measure, ncol = 2) +
  scale_y_log10(labels = label_number(accuracy = 0.0001)) +
  scale_color_brewer(palette = "Dark2", name = NULL) +
  labs(title = "Regional monthly percent of patients with a new diagnosis, log scale",
       subtitle = "Lyme disease, babesiosis, malaria and RMSF; same y-axis in every panel",
       x = NULL, y = "Percent of patients (log scale)") +
  theme_bw(base_size = 11) +
  theme(strip.text = element_text(face = "bold"),
        axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "bottom")
```

```{r nc-maps, eval=HAVE_NC, results='asis'}
nc_flag_cols <- paste0("epic_new_", DISEASES, "_suppressed_flag")

nc_state_annual <- nc %>%
  filter(nchar(geography) == 2, geography != "00") %>%
  mutate(year = substr(time, 1, 4)) %>%
  filter(year %in% YEARS) %>%
  group_by(geography, year) %>%
  summarize(across(all_of(c(paste0("epic_pct_new_", DISEASES),
                             paste0("epic_rate100k_new_", DISEASES))),
                    ~ mean(.x, na.rm = TRUE)),
            # TRUE only when every contributing month was suppressed
            across(all_of(nc_flag_cols), ~ mean(.x) == 1, .names = "allsupp_{.col}"),
            .groups = "drop") %>%
  mutate(year = factor(year, levels = YEARS))

draw_nc_map <- function(col, flag_col, label) {
  md <- states_sf %>%
    left_join(
      nc_state_annual %>%
        select(geography, year, value = all_of(col), suppressed = all_of(flag_col)) %>%
        mutate(value = if_else(suppressed, NA_real_, value)),
      by = "geography"
    ) %>%
    st_as_sf()

  ggplot(md) +
    geom_sf(aes(fill = value), color = "white", linewidth = 0.08) +
    scale_fill_viridis_c(option = "viridis", na.value = "grey70",
                         labels = label_number(accuracy = 0.01), name = NULL) +
    facet_wrap(~year, ncol = 2) +
    coord_sf(crs = 5070, datum = NA) +
    labs(title = label) +
    theme_void(base_size = 11) +
    theme(plot.title = element_text(face = "bold", hjust = 0.5, size = 13),
          strip.text = element_text(face = "bold", size = 11),
          legend.position = "right")
}

for (dz in DISEASES) {
  flag_col <- paste0("allsupp_epic_new_", dz, "_suppressed_flag")
  print(draw_nc_map(paste0("epic_pct_new_", dz), flag_col,
                    sprintf("%s — percent of patients with a new diagnosis, by year", DISEASE_LABELS[dz])))
  print(draw_nc_map(paste0("epic_rate100k_new_", dz), flag_col,
                    sprintf("%s — new diagnoses per 100,000 patients, by year", DISEASE_LABELS[dz])))
}
```

## Comparison with NNDSS and ArboNET

Epic Cosmos counts are compared with the NNDSS and `arbonet`. NNDSS has babesiosis, 
malaria, West Nile virus and dengue. ArboNET has West Nile virus only.

```{r external-data}
INGEST_RAW <- "https://raw.githubusercontent.com/PopHIVE/Ingest/main/data"

fetch_ingest <- function(path) {
  dest <- tempfile(fileext = ".csv.gz")
  ok <- tryCatch({
    download.file(file.path(INGEST_RAW, path), dest, mode = "wb", quiet = TRUE)
    TRUE
  }, error = function(e) FALSE)
  if (!ok) return(NULL)
  vroom(dest, show_col_types = FALSE,
        col_types = cols(geography = "c", time = "c"))
}

nnds_raw <- fetch_ingest("nnds/standard/data.csv.gz")
arbo_raw <- fetch_ingest("arbonet/standard/data_state.csv.gz")
HAVE_EXT <- !is.null(nnds_raw) && !is.null(arbo_raw)
if (!HAVE_EXT) message("Could not download NNDSS/ArboNET from the Ingest repo; skipping comparison.")
```

```{r external-annual, eval=HAVE_EXT}
SRC_EPIC <- "Epic Cosmos"; SRC_NNDS <- "NNDSS"; SRC_ARBO <- "ArboNET"
SRC_EPIC_NEW <- "Epic Cosmos (new cases)"
CMP_DISEASES <- c("babesiosis", "malaria", "west_nile", "dengue")
SRC_COLS <- c("Epic Cosmos" = "#1b6ca8", "Epic Cosmos (new cases)" = "#7b3294",
              "NNDSS" = "#e07b39", "ArboNET" = "#4c9a2a")

# Epic: annual patient counts from complete calendar years only. `prefix` is
# the numerator column prefix: epic_n_ (any diagnosis) or epic_n_new_ (first-time
# diagnoses, cosmos_vector_borne_new_cases).
annual_epic <- function(d, prefix, source) {
  d %>%
    mutate(year = as.integer(substr(time, 1, 4))) %>%
    group_by(geography, year) %>%
    summarize(across(all_of(paste0(prefix, DISEASES)), ~ sum(.x, na.rm = TRUE)),
              n_months = n(), .groups = "drop") %>%
    filter(n_months == 12) %>%
    pivot_longer(starts_with(prefix), names_to = "disease", values_to = "value") %>%
    mutate(disease = sub(paste0("^", prefix), "", disease), source = source) %>%
    select(geography, year, disease, source, value)
}

epic_annual <- annual_epic(dat, "epic_n_", SRC_EPIC)
if (HAVE_NC) epic_annual <- bind_rows(epic_annual, annual_epic(nc, "epic_n_new_", SRC_EPIC_NEW))

# NNDSS: values are cumulative year-to-date by week, so the annual total is the
# maximum over the year's weeks. Dengue = dengue + dengue-like + severe dengue.
NNDS_COLS <- list(
  babesiosis = "babesiosis",
  malaria    = "malaria",
  west_nile  = "arboviral_diseases_west_nile_virus_disease",
  dengue     = c("dengue_virus_infections_dengue",
                 "dengue_virus_infections_dengue_like_illness",
                 "dengue_virus_infections_severe_dengue")
)

nnds_annual <- nnds_raw %>%
  select(geography, year = mmwr_year, all_of(unname(unlist(NNDS_COLS)))) %>%
  group_by(geography, year) %>%
  summarize(across(everything(), ~ max(.x, na.rm = TRUE)), .groups = "drop")
for (dz in names(NNDS_COLS)) {
  nnds_annual[[dz]] <- rowSums(nnds_annual[NNDS_COLS[[dz]]])
}
nnds_annual <- nnds_annual %>%
  select(geography, year, all_of(names(NNDS_COLS))) %>%
  pivot_longer(all_of(names(NNDS_COLS)), names_to = "disease", values_to = "value") %>%
  group_by(disease, year) %>%
  # a disease-year with a national total of 0 was not reported, not zero cases
  mutate(value = if (max(value[geography == "00"]) == 0) NA_real_ else value) %>%
  ungroup() %>%
  mutate(source = SRC_NNDS, year = as.integer(year)) %>%
  select(geography, year, disease, source, value)

# ArboNET: West Nile virus cases (annual)
arbo_annual <- arbo_raw %>%
  transmute(geography, year = as.integer(substr(time, 1, 4)),
            disease = "west_nile", source = SRC_ARBO,
            value = arbonet_wnv_cases)

ann <- bind_rows(epic_annual, nnds_annual, arbo_annual) %>%
  filter(disease %in% CMP_DISEASES, year >= 2022, year <= 2025)
```

```{r rate-data, eval=HAVE_EXT}
STATE_FIPS <- all_fips %>%
  filter(nchar(geography) == 2, state %in% c(state.abb, "DC")) %>%
  select(geography, abbr = state)

pop_lookup <- vroom("../../resources/pop_state.csv.gz", show_col_types = FALSE) %>%
  filter(age_level == "Total") %>%
  inner_join(all_fips %>% filter(nchar(geography) == 2), by = c("geography" = "geography_name"),
             suffix = c("_name", "")) %>%
  filter(geography %in% STATE_FIPS$geography) %>%
  select(geography, pop = popsize)
pop_lookup <- bind_rows(pop_lookup,
                        tibble(geography = "00", pop = sum(pop_lookup$pop)))

# Yearly denominator: all Epic patients with any encounter that year, by state
# (cosmos_chronic state_year).
epic_annual_patients <- vroom("../cosmos_chronic/standard/state_year.csv.gz",
                              show_col_types = FALSE,
                              col_types = cols(geography = "c", time = "c")) %>%
  filter(age == "Total") %>%
  transmute(geography, year = as.integer(substr(time, 1, 4)),
            epic_n_patients = n_patients_chronic)

rate_epic <- function(d, prefix, source) {
  d %>%
    mutate(year = as.integer(substr(time, 1, 4))) %>%
    select(-epic_n_patients) %>%
    group_by(geography, year) %>%
    # Yearly scale: numerator = cases summed over the 12 months
    summarize(across(all_of(paste0(prefix, DISEASES)), ~ sum(.x, na.rm = TRUE)),
              n_months = n(), .groups = "drop") %>%
    filter(n_months == 12) %>%
    inner_join(epic_annual_patients, by = c("geography", "year")) %>%
    pivot_longer(all_of(paste0(prefix, DISEASES)), names_to = "disease", values_to = "num") %>%
    transmute(geography, year, disease = sub(paste0("^", prefix), "", disease),
              source = source, num, den = epic_n_patients)
}

epic_rate_parts <- rate_epic(dat, "epic_n_", SRC_EPIC)
if (HAVE_NC) epic_rate_parts <- bind_rows(epic_rate_parts,
                                          rate_epic(nc, "epic_n_new_", SRC_EPIC_NEW))

rate_parts <- bind_rows(
  epic_rate_parts,
  ann %>%
    filter(!source %in% c(SRC_EPIC, SRC_EPIC_NEW)) %>%
    inner_join(pop_lookup, by = "geography") %>%
    transmute(geography, year, disease, source, num = value, den = pop)
) %>%
  filter(disease %in% CMP_DISEASES, year >= 2022, year <= 2025, !is.na(num))
```

```{r compare-national, eval=HAVE_EXT}
nat_cmp <- ann %>%
  filter(geography == "00", !is.na(value)) %>%
  mutate(disease_label = factor(DISEASE_LABELS[disease],
                                levels = DISEASE_LABELS[CMP_DISEASES]))

ggplot(nat_cmp, aes(year, value, color = source)) +
  geom_line(linewidth = 0.7) +
  geom_point(size = 1.5) +
  facet_wrap(~disease_label, scales = "free_y", ncol = 2) +
  scale_y_log10(labels = label_comma()) +
  scale_color_manual(values = SRC_COLS, name = NULL) +
  scale_x_continuous(breaks = 2022:2025) +
  labs(title = "National annual counts",
       subtitle = "Epic Cosmos: patient counts; NNDSS and ArboNET: reported cases (log scale)",
       x = NULL, y = "Count (log scale)") +
  theme_bw(base_size = 11) +
  theme(strip.text = element_text(face = "bold"), legend.position = "bottom")

```

```{r rate-national, eval=HAVE_EXT}
nat_rate <- rate_parts %>%
  filter(geography == "00") %>%
  mutate(rate = num / den * 1e5,
         disease_label = factor(DISEASE_LABELS[disease],
                                levels = DISEASE_LABELS[CMP_DISEASES]))

ggplot(nat_rate, aes(year, rate, color = source)) +
  geom_line(linewidth = 0.7) +
  geom_point(size = 1.5) +
  facet_wrap(~disease_label, scales = "free_y", ncol = 2) +
  scale_y_log10(labels = label_number(accuracy = 0.01)) +
  scale_color_manual(values = SRC_COLS, name = NULL) +
  scale_x_continuous(breaks = 2022:2025) +
  labs(title = "National annual rate per 100,000",
       subtitle = "Epic Cosmos: per 100,000 Epic patients; NNDSS and ArboNET: per 100,000 US population (log scale)",
       x = NULL, y = "Rate per 100,000 (log scale)") +
  theme_bw(base_size = 11) +
  theme(strip.text = element_text(face = "bold"), legend.position = "bottom")

```

```{r rate-national-pct, eval=HAVE_EXT}
# National, no aggregation: Epic monthly, NNDSS weekly, ArboNET annual, each as
# a percent (Epic: of Epic patients that month; NNDSS/ArboNET: of US population).
US_POP <- pop_lookup$pop[pop_lookup$geography == "00"]

# Monthly denominator: that month's own patient count (epic_n_patients).
epic_pct_ts <- dat %>%
  filter(geography == "00") %>%
  mutate(time = as.Date(time)) %>%
  select(time, epic_n_patients, all_of(paste0("epic_n_", CMP_DISEASES))) %>%
  pivot_longer(starts_with("epic_n_") & !any_of("epic_n_patients"),
               names_to = "disease", values_to = "n") %>%
  transmute(time, disease = sub("^epic_n_", "", disease), source = SRC_EPIC,
            pct = n / epic_n_patients * 100)

# NNDSS values are cumulative year-to-date, so difference within each year to
# get the cases newly reported that week (dengue = the three dengue columns).
nnds_pct_ts <- nnds_raw %>%
  filter(geography == "00") %>%
  arrange(mmwr_year, mmwr_week) %>%
  transmute(time = as.Date(time), mmwr_year,
            babesiosis = babesiosis, malaria = malaria,
            west_nile = arboviral_diseases_west_nile_virus_disease,
            dengue = dengue_virus_infections_dengue +
              dengue_virus_infections_dengue_like_illness +
              dengue_virus_infections_severe_dengue) %>%
  pivot_longer(all_of(CMP_DISEASES), names_to = "disease", values_to = "cum") %>%
  group_by(disease, mmwr_year) %>%
  mutate(n = cum - lag(cum, default = 0)) %>%
  ungroup() %>%
  transmute(time, disease, source = SRC_NNDS, pct = n / US_POP * 100)

arbo_pct_ts <- arbo_raw %>%
  filter(geography == "00") %>%
  transmute(time = as.Date(paste0(substr(time, 1, 4), "-07-01")),
            disease = "west_nile", source = SRC_ARBO,
            pct = arbonet_wnv_cases / US_POP * 100)

epic_new_pct_ts <- if (HAVE_NC) {
  nc %>%
    filter(geography == "00") %>%
    mutate(time = as.Date(time)) %>%
    select(time, epic_n_patients, all_of(paste0("epic_n_new_", CMP_DISEASES))) %>%
    pivot_longer(starts_with("epic_n_new_"), names_to = "disease", values_to = "n") %>%
    transmute(time, disease = sub("^epic_n_new_", "", disease), source = SRC_EPIC_NEW,
              pct = n / epic_n_patients * 100)
}

pct_ts <- bind_rows(epic_pct_ts, epic_new_pct_ts, nnds_pct_ts, arbo_pct_ts) %>%
  filter(time >= as.Date("2022-01-01"), time < as.Date("2026-01-01"),
         !is.na(pct), pct > 0) %>%   # zeros / negative corrections can't be drawn on a log axis
  mutate(disease_label = factor(DISEASE_LABELS[disease], levels = DISEASE_LABELS[CMP_DISEASES]))

ggplot(pct_ts, aes(time, pct, color = source)) +
  geom_line(data = ~ filter(.x, source != SRC_ARBO), linewidth = 0.5) +
  geom_point(size = 0.9) +
  facet_wrap(~disease_label, scales = "free_y", ncol = 2) +
  scale_y_log10(labels = label_number(accuracy = 0.00001)) +
  scale_color_manual(values = SRC_COLS, name = NULL) +
  labs(title = "National percent of population, at each source's native resolution",
       subtitle = "Epic Cosmos: monthly, % of that month's Epic patients; NNDSS: weekly new cases, % of US population; ArboNET: annual (plotted mid-year), % of US population",
       x = NULL, y = "Percent (log scale)") +
  theme_bw(base_size = 11) +
  theme(strip.text = element_text(face = "bold"), legend.position = "bottom")
```

```{r compare-state, eval=HAVE_EXT}
# State share of the national total, pooled over the years (2023-2025) for
# which both sources have a national value.
pair_shares <- function(dz, src_x, src_y) {
  d <- ann %>% filter(disease == dz, source %in% c(src_x, src_y), year >= 2023)
  ok_years <- d %>%
    filter(geography == "00", !is.na(value)) %>%
    count(year) %>% filter(n == 2) %>% pull(year)
  d %>%
    filter(geography %in% STATE_FIPS$geography, year %in% ok_years) %>%
    group_by(source, geography) %>%
    summarize(value = sum(value, na.rm = TRUE), .groups = "drop") %>%
    group_by(source) %>%
    mutate(share = value / sum(value) * 100) %>%
    ungroup() %>%
    select(geography, source, share) %>%
    pivot_wider(names_from = source, values_from = share) %>%
    transmute(geography, x = .data[[src_x]], y = .data[[src_y]],
              disease = dz,
              panel = sprintf("%s (y) vs %s (x)", src_y, src_x),
              years = if (length(ok_years)) paste(range(ok_years), collapse = "-") else NA_character_)
}

plot_pairs <- function(pairs, facet_var, title, xlab, ylab) {
  cors <- pairs %>%
    group_by(.data[[facet_var]]) %>%
    summarize(label = sprintf("Spearman rho = %.2f\nyears %s",
                              cor(x, y, method = "spearman", use = "complete.obs"),
                              first(years)), .groups = "drop")
  pairs %>%
    left_join(STATE_FIPS, by = "geography") %>%
    ggplot(aes(x, y)) +
    geom_abline(linetype = "dashed", color = "grey60") +
    geom_point(color = "#1b6ca8", alpha = 0.7) +
    geom_text(aes(label = abbr), size = 2.6, vjust = -0.7, check_overlap = TRUE) +
    geom_text(data = cors, aes(x = -Inf, y = Inf, label = label),
              hjust = -0.05, vjust = 1.3, size = 3.2, inherit.aes = FALSE) +
    facet_wrap(vars(.data[[facet_var]]), scales = "free", ncol = 2) +
    labs(title = title, x = xlab, y = ylab,
         subtitle = "Each point is a state; dashed line = equal share") +
    theme_bw(base_size = 11) +
    theme(strip.text = element_text(face = "bold"))
}

epic_nnds <- bind_rows(lapply(CMP_DISEASES, pair_shares, SRC_EPIC, SRC_NNDS)) %>%
  mutate(disease = factor(DISEASE_LABELS[disease], levels = DISEASE_LABELS[CMP_DISEASES]))

plot_pairs(epic_nnds, "disease",
           "State share of national total: NNDSS vs. Epic Cosmos",
           "Epic Cosmos: % of national patient count",
           "NNDSS: % of national case count")

if (HAVE_NC) {
  epic_new_nnds <- bind_rows(lapply(CMP_DISEASES, pair_shares, SRC_EPIC_NEW, SRC_NNDS)) %>%
    mutate(disease = factor(DISEASE_LABELS[disease], levels = DISEASE_LABELS[CMP_DISEASES]))

  print(plot_pairs(epic_new_nnds, "disease",
                   "State share of national total: NNDSS vs. Epic Cosmos (new cases)",
                   "Epic Cosmos new cases: % of national patient count",
                   "NNDSS: % of national case count"))
}

wnv_pairs <- bind_rows(
  pair_shares("west_nile", SRC_EPIC, SRC_ARBO),
  if (HAVE_NC) pair_shares("west_nile", SRC_EPIC_NEW, SRC_ARBO),
  pair_shares("west_nile", SRC_NNDS, SRC_ARBO)
)

plot_pairs(wnv_pairs, "panel",
           "West Nile virus: state share of national total, ArboNET vs. other sources",
           "Other source: % of national count",
           "ArboNET: % of national case count")
```

```{r rate-state, eval=HAVE_EXT}
# State rate pooled over the years both sources have a national value
pair_rates <- function(dz, src_x, src_y) {
  d <- rate_parts %>% filter(disease == dz, source %in% c(src_x, src_y), year >= 2023)
  ok_years <- d %>% filter(geography == "00") %>% count(year) %>%
    filter(n == 2) %>% pull(year)
  d %>%
    filter(geography %in% STATE_FIPS$geography, year %in% ok_years) %>%
    group_by(source, geography) %>%
    summarize(rate = sum(num) / sum(den) * 1e5, .groups = "drop") %>%
    pivot_wider(names_from = source, values_from = rate) %>%
    transmute(geography, x = .data[[src_x]], y = .data[[src_y]],
              disease = dz,
              panel = sprintf("%s (y) vs %s (x)", src_y, src_x),
              years = if (length(ok_years)) paste(range(ok_years), collapse = "-") else NA_character_)
}

plot_rate_pairs <- function(pairs, facet_var, title, xlab, ylab) {
  cors <- pairs %>%
    group_by(.data[[facet_var]]) %>%
    summarize(label = sprintf("Spearman rho = %.2f\nyears %s",
                              cor(x, y, method = "spearman", use = "complete.obs"),
                              first(years)), .groups = "drop")
  pairs %>%
    left_join(STATE_FIPS, by = "geography") %>%
    ggplot(aes(x, y)) +
    geom_point(color = "#1b6ca8", alpha = 0.7) +
    geom_text(aes(label = abbr), size = 2.6, vjust = -0.7, check_overlap = TRUE) +
    geom_text(data = cors, aes(x = -Inf, y = Inf, label = label),
              hjust = -0.05, vjust = 1.3, size = 3.2, inherit.aes = FALSE) +
    facet_wrap(vars(.data[[facet_var]]), scales = "free", ncol = 2) +
    scale_x_continuous(trans = pseudo_log_trans(base = 10)) +
    scale_y_continuous(trans = pseudo_log_trans(base = 10)) +
    labs(title = title, x = xlab, y = ylab) +
    theme_bw(base_size = 11) +
    theme(strip.text = element_text(face = "bold"))
}

rate_epic_nnds <- bind_rows(lapply(CMP_DISEASES, pair_rates, SRC_EPIC, SRC_NNDS)) %>%
  mutate(disease = factor(DISEASE_LABELS[disease], levels = DISEASE_LABELS[CMP_DISEASES]))

plot_rate_pairs(rate_epic_nnds, "disease",
                "State rate: NNDSS (per 100,000 population) vs. Epic Cosmos (per 100,000 patients)",
                "Epic Cosmos: annual cases per 100,000 Epic patients",
                "NNDSS: cases per 100,000 population")

if (HAVE_NC) {
  rate_epic_new_nnds <- bind_rows(lapply(CMP_DISEASES, pair_rates, SRC_EPIC_NEW, SRC_NNDS)) %>%
    mutate(disease = factor(DISEASE_LABELS[disease], levels = DISEASE_LABELS[CMP_DISEASES]))

  print(plot_rate_pairs(rate_epic_new_nnds, "disease",
                        "State rate: NNDSS (per 100,000 population) vs. Epic Cosmos new cases (per 100,000 patients)",
                        "Epic Cosmos new cases: annual cases per 100,000 Epic patients",
                        "NNDSS: cases per 100,000 population"))
}

rate_wnv <- bind_rows(
  pair_rates("west_nile", SRC_EPIC, SRC_ARBO),
  if (HAVE_NC) pair_rates("west_nile", SRC_EPIC_NEW, SRC_ARBO),
  pair_rates("west_nile", SRC_NNDS, SRC_ARBO)
)

plot_rate_pairs(rate_wnv, "panel",
                "West Nile virus state rate",
                "Annual rate (Epic per 100,000 patients; NNDSS per 100,000 population)",
                "ArboNET: cases per 100,000 population")
```
