---
title: "PopHIVE Hepatitis C Care Cascade"
format:
  html:
    embed-resources: true
    toc: true
    toc-title: "On this page"
    toc-expand: true
    code-fold: true
    fig-align: center
    smart: false
execute:
  echo: false
  warning: false
  message: false
---

```{=html}
<link rel="preconnect" href="https://fonts.googleapis.com">
<link rel="preconnect" href="https://fonts.gstatic.com" crossorigin>
<link href="https://fonts.googleapis.com/css2?family=Rubik:wght@400;500;600&family=Source+Serif+4:opsz,wght@8..60,400;8..60,600&display=swap" rel="stylesheet">
<style>
body { font-family: 'Rubik', sans-serif; color: #26323d; background-color: #ffffff; padding-top: 2.75rem; }
h1, h2, h3, h4, .title, .subtitle { font-family: 'Source Serif 4', serif; color: #00356b; font-weight: 600; }
h2 { border-bottom: 2px solid #bedeff; padding-bottom: 0.3rem; margin-top: 2.5rem; }
a { color: #00356b; }
a:hover { color: #4e98e4; }
.quarto-title-banner .title, .quarto-title-banner .subtitle, .quarto-title-banner .quarto-title-meta { color: #ffffff !important; }
.quarto-title-banner { background-color: #00356b; }
::selection { background: #bedeff; }
.pophive-nav {
  position: fixed; top: 0; left: 0; right: 0; z-index: 1000;
  background: #00356b; padding: 0.7rem 1.5rem; display: flex; align-items: center;
  justify-content: space-between; font-family: 'Rubik', sans-serif;
}
.pophive-nav a { color: #ffffff !important; text-decoration: none; font-size: 0.95rem; font-weight: 500; }
.pophive-nav a:hover { color: #bedeff !important; }
.pophive-brand {
  font-family: 'Source Serif 4', serif !important; font-weight: 600 !important;
  font-size: 1.15rem !important; letter-spacing: 0.02em;
}
.pophive-nav-links { display: flex; gap: 1.75rem; }
.pophive-footer {
  width: 100vw; position: relative; left: 50%; right: 50%;
  margin-left: -50vw; margin-right: -50vw; margin-top: 3rem;
  padding: 1.75rem 1.5rem; text-align: center;
  background: #00356b; color: #dbe7f5; font-family: 'Rubik', sans-serif;
  font-size: 0.85rem; line-height: 1.6;
}
.pophive-footer strong { font-family: 'Source Serif 4', serif; color: #ffffff; font-size: 1rem; }
.pophive-footer a { color: #bedeff !important; }
.pophive-footer a:hover { color: #ffffff !important; }
.pophive-footer .disclaimer { color: #a9bdd4; font-size: 0.8rem; margin-top: 0.4rem; }
#TOC a { white-space: normal; word-wrap: break-word; }
.ctl-controls { display: flex; flex-wrap: wrap; align-items: center; gap: 1.25rem; margin: 0.25rem 0 0.5rem; font-family: 'Rubik', sans-serif; }
.ctl-field { font-size: 0.85rem; color: #00356b; font-weight: 500; }
.ctl-field select { font-family: 'Rubik', sans-serif; font-size: 0.85rem; color: #26323d; border: 1px solid #bedeff; border-radius: 6px; padding: 0.3rem 0.5rem; margin-left: 0.35rem; min-width: 12rem; }
.ctl-map-title { font-family: 'Source Serif 4', serif; color: #00356b; font-weight: 600; font-size: 1.15rem; margin: 0.5rem 0 0; }
.ctl-map-sub { font-family: 'Rubik', sans-serif; color: #767676; font-size: 0.82rem; margin: 0.15rem 0 0.25rem; }
.ctl-denom-note { font-family: 'Rubik', sans-serif; color: #767676; font-size: 0.78rem; font-style: italic; margin: 0 0 0.6rem; }
.ctl-slider-row { display: flex; align-items: center; gap: 0.75rem; margin: 0.25rem 0; font-family: 'Rubik', sans-serif; font-size: 0.82rem; color: #00356b; }
.ctl-slider-row input[type=range] { flex: 1; accent-color: #00356b; }
.ctl-period-label { min-width: 7rem; font-weight: 500; }
.overview-text { font-size: 14px; }
.state-picker-row { display: flex; align-items: center; gap: 0.75rem; margin: 1rem 0 1.5rem; padding: 0.85rem 1rem; background: #f5f9ff; border: 1px solid #bedeff; border-radius: 8px; }
.state-picker-row label { font-family: 'Source Serif 4', serif; font-weight: 600; color: #00356b; font-size: 1rem; }
.state-picker-row select { font-family: 'Rubik', sans-serif; font-size: 0.95rem; color: #26323d; border: 1px solid #bedeff; border-radius: 6px; padding: 0.35rem 0.6rem; min-width: 14rem; }
.suppressed-note { font-family: 'Rubik', sans-serif; color: #767676; font-size: 0.8rem; font-style: italic; margin-top: 0.4rem; }
.compare-toggle { display: flex; align-items: center; gap: 0.4rem; font-family: 'Rubik', sans-serif; font-size: 0.9rem; color: #00356b; margin-left: auto; }
.compare-toggle input { accent-color: #00356b; width: 1rem; height: 1rem; }
</style>
```

```{=html}
<nav class="pophive-nav">
  <a href="https://www.pophive.org/" class="pophive-brand">PopHIVE</a>
  <div class="pophive-nav-links">
    <a href="https://www.pophive.org/">Home</a>
    <a href="https://www.pophive.org/about">About</a>
  </div>
</nav>
```

## Overview

::: {.overview-text}
Notes:

- **Source**: Epic Cosmos, accessed via SlicerDicer
- **Restrictions**: Results of research performed with Epic Cosmos were obtained from the PopHIVE platform (https://github.com/PopHIVE/Ingest).
- **Suppression**: counts of 10 or fewer patients are suppressed and imputed as 5; suppressed years are shown as gaps in the trend charts and as a placeholder message in the care-cascade view.
:::

```{r}
#| label: setup
#| include: false
library(readr)
library(dplyr)
library(tidyr)
library(jsonlite)
```

```{r}
#| label: prep-data
#| include: false

data <- read_csv(
  "standard/data_dashboard.csv.gz", show_col_types = FALSE,
  col_types = cols(geography = col_character(), time = col_character())
)

state_lookup <- tribble(
  ~geography, ~state_name,
  "00", "National",
  "01", "Alabama",
  "17", "Illinois",
  "22", "Louisiana",
  "25", "Massachusetts"
)

records <- data %>%
  left_join(state_lookup, by = "geography") %>%
  mutate(
    year        = as.integer(substr(time, 1, 4)),
    partial     = !grepl("-12-31$", time),
    year_label  = if_else(partial, paste0(year, " (partial)"), as.character(year))
  ) %>%
  arrange(state_name != "National", state_name, time) %>%
  transmute(
    geography, state_name, time, year, year_label, partial,
    n                         = epic_n_patients_hepc_new_diagnosis,
    n_suppressed              = epic_n_patients_hepc_new_diagnosis_suppressed_flag,
    pct_meds                  = epic_pct_hepc_meds_after_diagnosis,
    n_meds                    = epic_n_patients_hepc_meds_after_diagnosis,
    pct_abnormal              = epic_pct_hepc_abnormal_test_after_diagnosis,
    n_abnormal                = epic_n_patients_hepc_abnormal_test_after_diagnosis,
    pct_any                   = epic_pct_hepc_any_test_after_diagnosis,
    pct_normal_after_meds     = epic_pct_normal_vl_20wk_after_meds_new_diagnosis,
    pct_normal_after_abnormal = epic_pct_normal_vl_20wk_after_abnormal_new_diagnosis
  )

states <- unique(records$state_name)
years  <- sort(unique(records$year))
```

```{r}
#| label: shared-data-json
#| results: asis

cat(paste0(
  '<script>',
  'window.HEPC_DATA = ', toJSON(records, na = "null", auto_unbox = TRUE), ';',
  'window.HEPC_STATES = ', toJSON(states, auto_unbox = TRUE), ';',
  'window.HEPC_YEARS = ', toJSON(years, auto_unbox = TRUE), ';',
  '</script>\n'
))
```

```{r}
#| label: plotly-dependency
# Renders nothing visible - registers the plotly.js htmlwidgets dependency so
# embed-resources: true inlines it. Every chart below is drawn by hand with
# vanilla JS against window.Plotly, matching the plain HTML controls used
# throughout.
htmltools::div(style = "display:none;", plotly::plot_ly())
```

```{=html}
<div class="state-picker-row">
  <label for="hepc-state">State:</label>
  <select id="hepc-state"></select>
  <label class="compare-toggle">
    <input type="checkbox" id="hepc-compare-national">
    Compare to National
  </label>
</div>
```

## Trends over time

Click a measure in the legend to show or hide it. A suppressed years appear as a gap in the line.

```{=html}
<p class="ctl-map-title" id="trend-title"></p>
<p class="ctl-map-sub">Percent of newly diagnosed hepatitis C patients, by outcome and year</p>
<div id="trend-chart" style="width:100%;height:480px;"></div>

<p class="ctl-map-title" id="count-title"></p>
<p class="ctl-map-sub">Number of patients newly diagnosed with hepatitis C, by year</p>
<div id="count-chart" style="width:100%;height:260px;"></div>

<div id="nat-count-wrap" style="display:none;">
<p class="ctl-map-title">National - newly diagnosed hepatitis C patients</p>
<p class="ctl-map-sub">Shown on its own scale - National's raw count is far larger than any one state's</p>
<div id="nat-count-chart" style="width:100%;height:260px;"></div>
</div>
```

```{=html}
<script>
(function () {
  function ready() {
    if (!window.Plotly) { return setTimeout(ready, 120); }
    var D = window.HEPC_DATA;
    var stateSel = document.getElementById('hepc-state');
    var compareChk = document.getElementById('hepc-compare-national');
    // National stays out of the picker (it's offered as a comparison overlay
    // instead, via the checkbox) -- only the 4 states with their own row.
    var PICKABLE_STATES = window.HEPC_STATES.filter(function (s) { return s !== 'National'; });
    PICKABLE_STATES.forEach(function (s) {
      var o = document.createElement('option'); o.value = s; o.textContent = s;
      stateSel.appendChild(o);
    });
    stateSel.value = PICKABLE_STATES[0];

    var byState = {};
    D.forEach(function (r) {
      (byState[r.state_name] = byState[r.state_name] || []).push(r);
    });

    // Categorical palette, fixed order (never re-cycled per selection) -
    // see the dataviz skill's validated default palette.
    var MEASURES = [
      { key: 'pct_meds',                  label: 'Received antivirals after diagnosis',        color: '#2a78d6' },
      { key: 'pct_abnormal',               label: 'Positive HCVL test after diagnosis',         color: '#eb6834' },
      { key: 'pct_any',                    label: 'Any HCVL test after diagnosis',              color: '#1baf7a' },
      { key: 'pct_normal_after_meds',      label: 'Normal VL ≥20wk after antivirals',        color: '#eda100' },
      { key: 'pct_normal_after_abnormal',  label: 'Normal VL ≥20wk after positive HCVL test', color: '#e87ba4' }
    ];

    var trendDiv = document.getElementById('trend-chart');
    var countDiv = document.getElementById('count-chart');
    var natCountDiv = document.getElementById('nat-count-chart');
    var natCountWrap = document.getElementById('nat-count-wrap');
    var trendTitle = document.getElementById('trend-title');
    var countTitle = document.getElementById('count-title');

    function render() {
      var rows = byState[stateSel.value] || [];
      var natRows = byState['National'] || [];
      var x = rows.map(function (r) { return r.year_label; });
      var compare = compareChk.checked;

      var traces = [];
      MEASURES.forEach(function (m) {
        traces.push({
          type: 'scatter', mode: 'lines+markers', name: m.label,
          x: x, y: rows.map(function (r) { return r[m.key]; }),
          line: { color: m.color, width: 2 }, marker: { size: 6, color: m.color },
          legendgroup: m.key, connectgaps: false,
          hovertemplate: m.label + '<br>%{x}<br>%{y:.1f}%<extra></extra>'
        });
        if (compare) {
          // National overlay: same color, dashed and lighter, sharing the
          // measure's legend group so toggling one hides both together.
          traces.push({
            type: 'scatter', mode: 'lines', name: m.label + ' (National)',
            x: natRows.map(function (r) { return r.year_label; }),
            y: natRows.map(function (r) { return r[m.key]; }),
            line: { color: m.color, width: 2, dash: 'dot' }, opacity: 0.55,
            legendgroup: m.key, connectgaps: false,
            hovertemplate: m.label + ' (National)<br>%{x}<br>%{y:.1f}%<extra></extra>'
          });
        }
      });

      Plotly.react(trendDiv, traces, {
        margin: { t: 20, b: 45, l: 55, r: 20 }, paper_bgcolor: 'white', plot_bgcolor: 'white',
        font: { family: 'Rubik, sans-serif', color: '#00356b', size: 11 }, hovermode: 'closest',
        xaxis: { title: 'Year', gridcolor: 'white', zeroline: false },
        yaxis: { title: 'Percent', range: [0, 100], gridcolor: '#e5e5e5', griddash: 'dash', zeroline: false },
        legend: { orientation: 'h', x: 0.5, xanchor: 'center', y: 1.22, yanchor: 'bottom' }
      }, { displayModeBar: false, responsive: true });

      var barColors = rows.map(function (r) { return r.n_suppressed === 1 ? '#c3c2b7' : '#2a78d6'; });
      var countTrace = {
        type: 'bar', name: stateSel.value, x: x, y: rows.map(function (r) { return r.n_suppressed === 1 ? null : r.n; }),
        marker: { color: barColors }, showlegend: false,
        hovertemplate: 'New diagnoses<br>%{x}<br>%{y:,}<extra></extra>'
      };
      // Suppressed years still get a visible placeholder bar (a thin sliver) so the
      // gap in the series reads as "unknown", not silently absent from the x-axis.
      var suppTrace = {
        type: 'bar', x: rows.filter(function (r) { return r.n_suppressed === 1; }).map(function (r) { return r.year_label; }),
        y: rows.filter(function (r) { return r.n_suppressed === 1; }).map(function () { return 1; }),
        marker: { color: '#c3c2b7' }, showlegend: false,
        hovertemplate: 'Suppressed (10 or fewer)<br>%{x}<extra></extra>'
      };
      // National's raw patient count is 100-1000x a single state's, so it is
      // deliberately NOT overlaid here (that would need a second y-axis, or
      // it would flatten the state's own bars to invisible) -- the National
      // comparison for counts lives in its own chart below instead.
      Plotly.react(countDiv, [countTrace, suppTrace], {
        barmode: 'overlay',
        margin: { t: 10, b: 40, l: 60, r: 20 }, paper_bgcolor: 'white', plot_bgcolor: 'white',
        font: { family: 'Rubik, sans-serif', color: '#00356b', size: 11 }, hovermode: 'closest',
        xaxis: { title: 'Year', gridcolor: 'white', zeroline: false },
        yaxis: { title: 'Patients', gridcolor: '#e5e5e5', griddash: 'dash', zeroline: false, rangemode: 'tozero' },
        showlegend: false
      }, { displayModeBar: false, responsive: true });

      if (compare) {
        var natBarColors = natRows.map(function (r) { return r.n_suppressed === 1 ? '#c3c2b7' : '#00356b'; });
        Plotly.react(natCountDiv, [{
          type: 'bar', x: natRows.map(function (r) { return r.year_label; }),
          y: natRows.map(function (r) { return r.n_suppressed === 1 ? null : r.n; }),
          marker: { color: natBarColors },
          hovertemplate: 'National new diagnoses<br>%{x}<br>%{y:,}<extra></extra>'
        }, {
          type: 'bar',
          x: natRows.filter(function (r) { return r.n_suppressed === 1; }).map(function (r) { return r.year_label; }),
          y: natRows.filter(function (r) { return r.n_suppressed === 1; }).map(function () { return 1; }),
          marker: { color: '#c3c2b7' },
          hovertemplate: 'Suppressed (10 or fewer)<br>%{x}<extra></extra>'
        }], {
          barmode: 'overlay',
          margin: { t: 10, b: 40, l: 60, r: 20 }, paper_bgcolor: 'white', plot_bgcolor: 'white',
          font: { family: 'Rubik, sans-serif', color: '#00356b', size: 11 }, hovermode: 'closest',
          xaxis: { title: 'Year', gridcolor: 'white', zeroline: false },
          yaxis: { title: 'Patients', gridcolor: '#e5e5e5', griddash: 'dash', zeroline: false, rangemode: 'tozero' },
          showlegend: false
        }, { displayModeBar: false, responsive: true });
      }
      natCountWrap.style.display = compare ? '' : 'none';

      trendTitle.textContent = stateSel.value + ' - hepatitis C care cascade rates' + (compare ? ' (dotted = National)' : '');
      countTitle.textContent = stateSel.value + ' - newly diagnosed hepatitis C patients';
    }

    stateSel.addEventListener('change', function () { render(); renderCascade(); });
    compareChk.addEventListener('change', function () { render(); renderCascade(); });
    render();
  }
  ready();
})();
</script>
```

## Care cascade

```{=html}
<div class="ctl-slider-row">
  <span class="ctl-period-label" id="cascade-year-label"></span>
  <input type="range" id="cascade-year" min="1" value="1" step="1" aria-label="year">
</div>
<p class="ctl-map-title" id="cascade-title"></p>
<p class="ctl-map-sub" id="cascade-sub"></p>
<div id="cascade-chart" style="width:100%;height:480px;"></div>
<p class="ctl-denom-note" id="cascade-note"></p>

<div id="nat-cascade-wrap" style="display:none;">
<p class="ctl-map-title" id="nat-cascade-title"></p>
<p class="ctl-map-sub" id="nat-cascade-sub"></p>
<div id="nat-cascade-chart" style="width:100%;height:480px;"></div>
</div>
```

```{=html}
<script>
(function () {
  function ready() {
    if (!window.Plotly) { return setTimeout(ready, 120); }
    var D = window.HEPC_DATA;
    var YEARS = window.HEPC_YEARS;
    var stateSel = document.getElementById('hepc-state');
    var compareChk = document.getElementById('hepc-compare-national');
    var slider = document.getElementById('cascade-year');
    var yearLabel = document.getElementById('cascade-year-label');
    var titleEl = document.getElementById('cascade-title');
    var subEl = document.getElementById('cascade-sub');
    var noteEl = document.getElementById('cascade-note');
    var chartDiv = document.getElementById('cascade-chart');
    var natWrap = document.getElementById('nat-cascade-wrap');
    var natTitleEl = document.getElementById('nat-cascade-title');
    var natSubEl = document.getElementById('nat-cascade-sub');
    var natChartDiv = document.getElementById('nat-cascade-chart');

    var byState = {};
    D.forEach(function (r) {
      (byState[r.state_name] = byState[r.state_name] || []).push(r);
    });

    slider.max = YEARS.length;
    // Default to the most recent FULL year (2026 is a partial, fully-suppressed
    // year in every geography), not the last slider position.
    var defaultIdx = YEARS.length;
    for (var i = YEARS.length - 1; i >= 0; i--) {
      var probe = (byState['National'] || []).filter(function (r) { return r.year === YEARS[i]; })[0];
      if (probe && probe.n_suppressed !== 1) { defaultIdx = i + 1; break; }
    }
    slider.value = defaultIdx;

    var fmt = function (n) { return Math.round(n).toLocaleString(); };

    // Draws one state/year's cascade into `div`; returns the subtitle text
    // (the caller decides where to put it). A true sequential cascade now:
    // stage-2 boxes (normal VL) are counted directly off their stage-1 box's
    // own count, not off the full new-diagnosis population - see the
    // Overview note on the 2026-09-21 recondition.
    function drawCascade(div, label, row) {
      if (!row || row.n_suppressed === 1) {
        var suppText = row && row.partial
          ? 'Data suppressed: ' + row.year + ' is a partial year with 10 or fewer new diagnoses in ' + label + '.'
          : 'Data suppressed for ' + label + ' (10 or fewer new diagnoses).';
        Plotly.react(div, [], {
          annotations: [{
            text: suppText, showarrow: false, x: 0.5, y: 0.5,
            xref: 'paper', yref: 'paper', font: { size: 13, color: '#767676' }
          }],
          margin: { t: 10, b: 10, l: 10, r: 10 }, paper_bgcolor: 'white', plot_bgcolor: 'white'
        }, { displayModeBar: false, responsive: true });
        return 'Newly diagnosed hepatitis C patients';
      }

      var n = row.n;
      var nMeds     = row.n_meds;
      var nAbnormal = row.n_abnormal;
      var nAny      = n * row.pct_any / 100;
      var nNormAfterMeds     = nMeds * row.pct_normal_after_meds / 100;
      var nNormAfterAbnormal = nAbnormal * row.pct_normal_after_abnormal / 100;

      // Short labels stay ON the diagram (name + count only, so six of them
      // can sit in three columns without their text runs colliding); the
      // full description + percent moves into customdata, read out on hover
      // instead - see the hovertemplate below.
      var labels = [
        'New diagnoses<br>' + fmt(n),
        'Antivirals<br>' + fmt(nMeds),
        'Positive test<br>' + fmt(nAbnormal),
        'Any test<br>' + fmt(nAny),
        'Normal VL<br>' + fmt(nNormAfterMeds),
        'Normal VL<br>' + fmt(nNormAfterAbnormal)
      ];
      var hoverLabels = [
        'New HCV diagnoses<br>' + fmt(n) + ' patients',
        'Received antivirals after diagnosis<br>' + fmt(nMeds) + ' patients (' + row.pct_meds.toFixed(1) + '% of new diagnoses)',
        'Positive HCVL test after diagnosis<br>' + fmt(nAbnormal) + ' patients (' + row.pct_abnormal.toFixed(1) + '% of new diagnoses)',
        'Any HCVL test after diagnosis<br>' + fmt(nAny) + ' patients (' + row.pct_any.toFixed(1) + '% of new diagnoses)',
        'Normal VL ≥20wk after antivirals<br>' + fmt(nNormAfterMeds) + ' patients (' + row.pct_normal_after_meds.toFixed(1) + '% of patients who received antivirals)',
        'Normal VL ≥20wk after a positive test<br>' + fmt(nNormAfterAbnormal) + ' patients (' + row.pct_normal_after_abnormal.toFixed(1) + '% of patients with a positive test)'
      ];
      var nodeColors = ['#00356b', '#2a78d6', '#eb6834', '#1baf7a', '#86b6ef', '#f2b880'];
      var nodeX = [0.01, 0.4, 0.4, 0.4, 0.98, 0.98];
      var nodeY = [0.5, 0.05, 0.5, 0.95, 0.05, 0.5];
      var linkColors = ['rgba(42,120,214,0.35)', 'rgba(235,104,52,0.35)', 'rgba(27,175,122,0.35)',
        'rgba(42,120,214,0.25)', 'rgba(235,104,52,0.25)'];

      var trace = {
        type: 'sankey', arrangement: 'snap',
        textfont: { family: 'Rubik, sans-serif', size: 11, color: '#26323d' },
        node: {
          label: labels, customdata: hoverLabels, color: nodeColors, pad: 28, thickness: 16,
          line: { color: 'white', width: 1 }, x: nodeX, y: nodeY,
          hovertemplate: '%{customdata}<extra></extra>'
        },
        link: {
          source: [0, 0, 0, 1, 2],
          target: [1, 2, 3, 4, 5],
          value:  [nMeds, nAbnormal, nAny, nNormAfterMeds, nNormAfterAbnormal],
          color:  linkColors,
          hovertemplate: '%{value:.0f} patients<extra></extra>'
        }
      };

      Plotly.react(div, [trace], {
        margin: { t: 10, b: 10, l: 10, r: 10 }, paper_bgcolor: 'white', plot_bgcolor: 'white',
        font: { family: 'Rubik, sans-serif', color: '#00356b', size: 12 }
      }, { displayModeBar: false, responsive: true });

      return 'Newly diagnosed hepatitis C patients (n = ' + fmt(n) + ')' + (row.partial ? ' - partial year' : '');
    }

    function render() {
      var rows = byState[stateSel.value] || [];
      var year = YEARS[parseInt(slider.value, 10) - 1];
      yearLabel.textContent = String(year);
      var row = rows.filter(function (r) { return r.year === year; })[0];
      var compare = compareChk.checked;

      titleEl.textContent = stateSel.value + ' - ' + year + ' care cascade';
      subEl.textContent = drawCascade(chartDiv, stateSel.value, row);
      noteEl.textContent = 'The two rightmost boxes are computed against the box immediately upstream (a true sequential cascade) - see the Overview note above.';

      natWrap.style.display = compare ? '' : 'none';
      if (compare) {
        var natRow = (byState['National'] || []).filter(function (r) { return r.year === year; })[0];
        natTitleEl.textContent = 'National - ' + year + ' care cascade';
        natSubEl.textContent = drawCascade(natChartDiv, 'National', natRow);
      }
    }

    slider.addEventListener('input', render);
    window.renderCascade = render;
    render();
    // The sankey can draw blank on this very first call, before the page's
    // layout (fonts, TOC sidebar) has settled - a known plotly.js quirk.
    // Re-render once more on the next frame and once more on window load,
    // the same fix that happens today when the slider is moved by hand.
    requestAnimationFrame(render);
    window.addEventListener('load', render);
  }
  ready();
})();
</script>
```

```{=html}
<nav class="pophive-footer">
  <strong>PopHIVE</strong><br>
  A public health surveillance platform combining multiple data sources.<br>
  <a href="https://www.pophive.org/about">About PopHIVE</a> &middot;
  <a href="https://github.com/PopHIVE/Ingest">Data Processing on GitHub</a>
  <div class="disclaimer">Data shown here comes from Epic Cosmos and is subject to the suppression and imputation rules described above. This page is for informational purposes and is not a substitute for clinical judgment.</div>
</nav>
```
